Community-Conditioned Language Models

### 2021-05-04 paper/paper.tex
restructure
some todos and outline of work

\begin{document}
\maketitle
\begin{abstract}
  % Context (Optional)

@@

% Data sets

% Community Conditioned Language Models:
%   Architecture (with emphasis on the embeddings)
%   Training
%   Results (loss and loss per community)

@@

% Correlation between linguistic embedding with social embeddings
%    Explain the problem (idea: comparing cosine similarities --> too much data, again)
%    Sketch "procrustes" problem, explain what it yields. (Also what kind of similarity one gets with random input)
%    Results:
%      - Correlation between Kumar2018 and best LSTM
%      - Correlation between Kumar2018 and best Transformer 
%      - Correlation between Transformer and LSTM
%      - Concatenations ???

@@

@@

\section{Results and analysis}

\subsection{Losses of CCLM}

@@

\subsection{Analysis of CCLM community embeddings}

As a qualitative assessment of the community embeddings, 
we manually identified topic categories into which at least three
of the selected communities fall.
We plotted the two-component PCA projection of the embeddings,
annotated with these categories (\cref{fig:comm-pca}).
We see that there is some evidence of clustering according to
the topic categories we identified,
suggesting that the embedding space learned by the model is meaningful.

\begin{figure*}[t]
\begin{tikzpicture}
  \begin{groupplot}[
    group style={group size=2 by 1},
    yticklabels={,,}, xticklabels={,,},
    legend style={anchor=north, legend columns=1, font=\sffamily\tiny},
  ]
    \nextgroupplot
    \commscatter{lstm-3-1_pca1}{lstm-3-1_pca2}
    \nextgroupplot
    \commscatter{transformer-3-3_pca1}{transformer-3-3_pca2}
    \legend{General interest,Videogames,Technology,Sports,Female-focused,Other}
  \end{groupplot}
\end{tikzpicture}
\caption{Community embedding PCA of an LSTM (left, $c=1$) and transformer (right, $c=3$) model.}
\label{fig:comm-pca}
\end{figure*}

As a quantitative assessment, we compare the CCLM-learned community
embedding with the community embedding created by \citet{Kumar2018},
who generated them using using a negative-sampling optimization
algorithm and the message-community co-occurrence matrix as ground
truth, using data which spans January 2014 to April 2017.
We refer the reader
to \citet{Kumar2018}
for details, but the important point is that no linguistic
information is used to create these embeddings: they only depend on
whether a given user participates in a given
community. In sharp contrasts, CCLM community embeddings depend in no
way on which user is the author of any given message: we only use the
contents of messages, not authorship data.
%
We correlate the pairwise cosine similarity of the
CCLM embeddings with the same metric on the membership embedding,
restricted to the 46 communities chosen for our study
(\cref{fig:pairwise-comm-sim}).

\begin{figure}
  \pairwisecommsim{snap_cos_sim}{lstm-3-1_cos_sim} 
  \caption{Pairwise cosine similarity between community vectors for 
    the membership-based embedding (x-axis) and
  the LSTM CCLM embedding with $c=1$ (y-axis). 
  Pairs from the same category are colored according to \cref{fig:comm-pca}.
  We also mark pairs from different categories (\texttimes) and pairs where one or both
  communities belong to the ``Other'' category (\textbigcircle).
  }
  \label{fig:pairwise-comm-sim}
\end{figure}

\begin{table}
  \small
  \centering
  \input{floats/comm_sim.tex}
  \caption{Pearson's $r$ between the pairwise similarity of community vectors
  in the CCLM models and the membership-based embedding of \citet{Kumar2018}.
  ($p<0.001$ for all models).
  }
  \label{tab:pairwise-comm-sim}
\end{table}

@@

>>>>

\begin{document}
\maketitle
\begin{abstract}

  TODO: fix this up.
  % Context (Optional)

@@

% Community Conditioned Language Models:
%   Data sets
%   Architecture (with emphasis on the embeddings)
%   Training
%   Results (loss and loss per community)

@@

% Correlation between linguistic embedding with social embeddings
%    Kumar2018 Data set
%    Explain the problem (idea: comparing cosine similarities --> too much data, again)
%    Sketch "procrustes" problem, explain what it yields. (Also what kind of similarity one gets with random input)
%    Results:
%      - Correlation between Kumar2018 and best LSTM
%      - Correlation between Kumar2018 and best Transformer 
%      - Correlation between Transformer and LSTM
%      - Concatenations ???

@@

TODO: outline of the paper.

@@

\subsection{Results and analysis}

\subsubsection{Losses of CCLM}

@@

@@

\section{Social Analysis of CCLM community embeddings (TODO: better title)}

- General idea; repeated.

\subsection{Data set}
We compare the CCLM-learned community
embedding with the community embedding created by \citet{Kumar2018},
who generated them using using a negative-sampling optimization
algorithm and the message-community co-occurrence matrix as ground
truth, using data which spans January 2014 to April 2017.
We refer the reader
to \citet{Kumar2018}
for details, but the important point is that no linguistic
information is used to create these embeddings: they only depend on
whether a given user participates in a given
community. In sharp contrasts, CCLM community embeddings depend in no
way on which user is the author of any given message: we only use the
contents of messages, not authorship data.
%

\subsection{Idea (JP)}
% Explain the problem (idea: comparing cosine similarities --> too much data, again)

correlate the pairwise cosine similarity of the
CCLM embeddings with the same metric on the membership embedding,
restricted to the 46 communities chosen for our study
(\cref{fig:pairwise-comm-sim}).

\subsection{Procrustes (JP)}

%    Sketch "procrustes" problem, explain what it yields. (Also what kind of similarity one gets with random input)

\subsection{Results}

%    Results:
%      - Correlation between Kumar2018 and best LSTM
%      - Correlation between Kumar2018 and best Transformer 
%      - Correlation between Transformer and LSTM
%      - Concatenations ???

\begin{figure*}[t]
\begin{tikzpicture}
  \begin{groupplot}[
    group style={group size=2 by 1},
    yticklabels={,,}, xticklabels={,,},
    legend style={anchor=north, legend columns=1, font=\sffamily\tiny},
  ]
    \nextgroupplot
    \commscatter{lstm-3-1_pca1}{lstm-3-1_pca2}
    \nextgroupplot
    \commscatter{transformer-3-3_pca1}{transformer-3-3_pca2}
    \legend{General interest,Videogames,Technology,Sports,Female-focused,Other}
  \end{groupplot}
\end{tikzpicture}
\caption{Plot of embeddings along the 2 principal dimensions of correlation between .... TODO of an LSTM (left, $c=1$) and transformer (right, $c=3$) model.
Color correspond the 5-means clustering of .... TODO}
\label{fig:comm-pca}
\end{figure*}
